---
title: "Workshop 2: Instructor Answer Key"
subtitle: "Entomology 3927 - Fall 2026"
author: Jeremy Hemberger
date: last-modified
date-format: long
engine: knitr
format:
  live-html:
    theme: 
    - ../styles.scss
    highlight-style: ayu
---

## Vectors
### Exercise 1
```{r}
species <- c("Apis mellifera", "Bombus impatiens", "Osmia lignaria")
individual_counts <- c(12, 45, 8)
survey_data <- c(species, individual_counts)
class(survey_data)
# [1] "character"
```

### Exercise 2
```{r}
dry_weight_mg <- c(120, 340, 85, 510, 230, 410)
dry_weight_g <- dry_weight_mg / 1000
net_mass_g <- dry_weight_g - c(0.005, 0.008)
net_mass_g
# [1] 0.115 0.332 0.080 0.502 0.225 0.402
```

### Exercise 3
```{r}
stem_density <- c(145, 212, 88, 304, 199, 250, 180, 218)
dense_even_plots <- stem_density[stem_density > 200 & stem_density %% 2 == 0]
dense_even_plots
# [1] 212 304 250 218
```

### Exercise 4
```{r}
soil_moisture <- c(0.24, NA, 0.31, 0.19, NA, 0.28, 0.22)
mean_moisture <- mean(soil_moisture, na.rm = TRUE)
soil_moisture[is.na(soil_moisture)] <- mean_moisture
soil_moisture
# [1] 0.24 0.248 0.31 0.19 0.248 0.28 0.22
```

### Exercise 5
```{r}
shannon_index <- c(Site_A = 1.82, Site_B = 2.45, Site_C = 1.31, Site_D = 2.89, Site_E = 2.10, Ctrl_1 = 0.45, Ctrl_2 = 0.62)
target_sites <- shannon_index[c("Site_B", "Site_D")]
intact_sites <- shannon_index[-c(6, 7)]

target_sites
# Site_B Site_D 
#   2.45   2.89 

intact_sites
# Site_A Site_B Site_C Site_D Site_E 
#   1.82   2.45   1.31   2.89   2.10
```


## Lists
### Exercise 1
```{r}
site_data <- list(
  site_id = "MN_BEE_01",
  species = c("Lasioglossum vierecki", "Bombus vagans", "Ceratina dupla", "Agapostemon virescens"),
  environmental = list(temperatures = c(21.5, 23.0, 22.8), humidity = 64)
)

temp_vals <- site_data$environmental$temperatures
third_species <- site_data$species[[3]]

temp_vals
# [1] 21.5 23.0 22.8

third_species
# [1] "Ceratina dupla"
```

### Exercise 2
```{r}
trap_counts <- list(
  Prairie = c(12, 18, 25, 9),
  Forest_Edge = c(5, 8, 3, 11, 4),
  Agriculture = c(30, 42, 38)
)

total_list <- lapply(trap_counts, sum)
site_totals <- unlist(total_list)

site_totals
#     Prairie Forest_Edge Agriculture 
#          64          31         110
```

### Exercise 3

```{r}
pollinator_obs <- list(
  bees = c(45, 32),
  flies = c(12, 19)
)

pollinator_obs$butterflies <- c(8, 14)
pollinator_obs$bees <- pollinator_obs$bees + 10

pollinator_obs
# $bees
# [1] 55 42
# 
# $flies
# [1] 12 19
# 
# $butterflies
# [1]  8 14
```

## Tibbles/data frames
### Exercise 1

```{r}
library(tibble)

tree_survey <- tibble(
  species = c("Quercus alba", "Acer saccharum", "Pinus strobus", "Quercus rubra", "Betula alleghaniensis"),
  dbh_cm = c(12.4, 28.1, 8.5, 42.0, 19.3),
  basal_area_m2 = (pi * (dbh_cm / 2)^2) / 10000
)

tree_survey
```

### Exercise 2

```{r}
df_bees <- data.frame(site = c("A", "B", "C"), parasite_load = c(0, 14, 3))
tbl_bees <- as_tibble(df_bees)

df_subset_class <- class(df_bees[, "parasite_load"])
tbl_subset_class <- class(tbl_bees[, "parasite_load"])

df_subset_class
# [1] "numeric"

tbl_subset_class
# [1] "tbl_df"     "tbl"        "data.frame"
```

### Exercise 3

```{r}
df_plots <- data.frame(treatment_group = c("Control", "Burned", "Grazed"), diversity = c(1.2, 2.4, 1.8))
tbl_plots <- as_tibble(df_plots)

df_match <- df_plots$treat
tbl_match <- tbl_plots$treat

df_match
# [1] "Control" "Burned"  "Grazed"

tbl_match
# NULL
# Warning message: Unknown or uninitialised column: `treat`.
```

### Exercise 4

```{r}
water_quality <- data.frame(
  site_id = paste0("Site_", 1:6),
  pH = c(6.8, 7.2, 7.1, 6.5, 7.4, 6.9),
  dissolved_oxygen = c(8.2, 6.4, 7.9, 5.8, 8.0, 7.1)
)

top_samples <- water_quality[1:3, ]
specific_do <- water_quality[4, "dissolved_oxygen"]
pH_sub <- water_quality[2:5, c("site_id", "pH")]

top_samples
specific_do
# [1] 5.8

pH_sub
```

### Exercise 5
```{r}
set.seed(42)
df_sweeps <- data.frame(
  sweep_id = 1:100,
  sweeps_count = sample(15:50, 100, replace = TRUE),
  biomass_g = runif(100, 0.5, 12.0)
)

tbl_sweeps <- as_tibble(df_sweeps)
print(tbl_sweeps)
# # A tibble: 100 × 3
#    sweep_id sweeps_count biomass_g
#       <int>        <int>     <dbl>
#  1        1           48      9.51
#  2        2           47     10.5 
#  3        3           25      5.70
#  4        4           45      3.38
#  5        5           38      1.80
#  6        6           33     11.1 
#  7        7           41      5.57
#  8        8           40      4.98
#  9        9           34      8.72
# 10       10           40      1.19
# # ℹ 90 more rows
```

## Code Style
Key Style Violations for Students to Fix

    * Variable Names: Replaced non-descriptive (T, x, T_2), camelCase (MeanCount, FinalData), and mixed Camel_Snake_Case (Filter_Data_fn) with clean snake_case. Overwrote the reserved logical T.

    * Spacing around operators: Added spaces around assignment (<-, =) and mathematical operators (+, -, /, %>%, ==).

    * Indentation and Line Breaks: Cleaned up crammed curly braces {} and functions squeezed onto single lines with semicolons ;.

    * Pipes & Long Lines: Split long tidyverse pipelines and ggplot layers across multiple lines (keeping lines under 80 characters) with operators placed at the end of lines.

    * Commas: Ensured a single space follows all commas (c(x, y) not c(x,y)).

```{r}
library(tidyverse)

# Load raw field sampling data
field_data <- read.csv("field_sampling_data.csv")

# Filter for active observations
active_data <- field_data[field_data$count > 0, ]

mean_count <- mean(active_data$count)

# Filter site data and calculate standardized counts
filter_site_data <- function(df, species_name, site_name) {
  if (species_name %in% df$species) {
    df_sub <- df[df$species == species_name & df$site == site_name, ]
    df_sub$count_scaled <- (df_sub$count - mean_count) / sd(df_sub$count)
    return(df_sub)
  } else {
    print("Species not found in this site!")
    return(NULL)
  }
}

final_data <- filter_site_data(active_data, "Apis mellifera", "Site_A")

# Summarize species totals
site_summary <- final_data %>%
  group_by(species, site) %>%
  summarize(
    total = sum(count),
    mean_scaled = mean(count_scaled),
    sd_scaled = sd(count_scaled)
  ) %>%
  filter(total > 10)

# Plot raw vs. scaled counts
ggplot(final_data, aes(x = count, 
                       y = count_scaled, 
                       color = species)) +
  geom_point() +
  theme_bw()
```